Very simplified likelihood

### 2024-02-23 inst/very_simplified_likelihood.Rmd
include Stirling

\[
\frac{(B-1)(B+n-r-2)!}{(B+n-1)!}
\]

@@

	if(log){
		out <- 0
		for(i in seq_len(nrow(M))){out <- out + f_vec_a(a,r=M[i,1],n=M[i,2],log=TRUE)}
	} else {
		out <- 1
		for(i in seq_len(nrow(M))){out <- out * f_vec_a(a,r=M[i,1],n=M[i,2],log=FALSE)}
	}
	return(out)
}
```

@@

```{r,label=defmle}
MLE <- function(r,n,give=FALSE){
    d <- 1e-6
     jj <- optimize(f_vec_arn,c(d,1-d),r=r,n=n,log=TRUE,maximum=TRUE)
	 if(!give){jj <- jj$maximum}
	 return(jj)
}

MLEa <- function(r,n,give=FALSE){
	jj <- uniroot(
		f=function(a){fdash(a,r,n)},
		interval=c(0.01,0.99))
	if(!give){jj <- jj$root}
	return(jj)
}

@@

>>>>

\[
\mathcal{L}_{r,n}(a)=\frac{(B-1)(B+n-r-2)!}{(B+n-1)!},\qquad B=1/(1-a)
\]

Then, using Stirling [viz $n!\sim n^ne^{-n}\sqrt{2\pi n}$] we get

\[
\log(B-1) + ((B+n-r-2)\log(B+n-r-2)-(B+n-r-2)+\frac{\log(B+n-r-2)}{2})
          - ((B+n  -1)\log(B+n  -1)-(B+n  -1)+\frac{\log(B+n  -1)}{2})
\]

If you give Sage this:

```
var('B n r')
pretty_print_default(True)
X=log(B-1) + (
+ ((B+n-r-2)*log(B+n-r-2)-(B+n-r-2)+log(B+n-r-2)/2)    
- ((B+n  -1)*log(B+n  -1)-(B+n  -1)+log(B+n  -1)/2)
)
X.taylor(n,Infinity,2)
```

It returns this:

\[
-(r+1)\log n
-\frac{(2B-3)r-r^2+2B-2}{2n}
-\frac{3(2B-3)r^2-6B^2-6(B^2-3B+2)r+12B-5}{12n^2}
+\log(B-1)+\mathcal{O}(n^{-3})
\]

After simplifying and taking an asymptotic expansion we find

\[
\mathcal{S}=\log\mathcal{L}_{r,n}(a)=\log(B-1)-B\frac{r+1}{n} + \frac{B^2-B(5+r^2)/2}{n^2}
+ K + \mathcal{O}\left(n^{-3}\right)
\]

(where $K$ is an arbitrary constant).  If we then neglect terms of
order $n^{-2}$ and above, viz $\log(B-1)-B\frac{r+1}{n}$, observing
that $\partial^2\mathcal{S}/\partial B^2=-(B-1)^{-2}<0$, the evaluate
would be $\hat{a}=\frac{n}{n+r+1}$ [although this seems to be a poor
approximation to the exact evaluate]

@@

	if(log){
		out <- 0
		for(i in seq_len(nrow(M))){out <- out + f_vec_a(a,r=M[i,1],n=M[i,2],log=TRUE)}
	} else {
		out <- 1
		for(i in seq_len(nrow(M))){out <- out * f_vec_a(a,r=M[i,1],n=M[i,2],log=FALSE)}
	}
	return(out)
}

fapprox <- function(a,r,n,log=FALSE){
	if(log){
		return(log(B-1)-B*(r+1)/n)
	} else {
		return(B-1 - (r+1)*exp(B)/n)
	}
}
```

@@

```{r,label=defmle}
MLE <- function(r,n,give=FALSE){
	d <- 1e-6
	out <- optimize(f_vec_arn,c(d,1-d),r=r,n=n,log=TRUE,maximum=TRUE)
	if(!give){out <- out$maximum}
	return(out)
}

MLEa <- function(r,n,give=FALSE){
	out <- uniroot(f=function(a){fdash(a,r,n)},interval=c(0.01,0.99))
	if(!give){out <- out$root}
	return(out)
}

@@

```{r usestirling}
rbind(
c(MLE(10 , 100),100 / 111),
c(MLE(100,1000),1000/1101),
c(MLE(500,5000),5000/5501)
)

rbind(
c(MLEa(10 , 100),100 / 111),
c(MLEa(100,1000),1000/1101),
c(MLEa(500,5000),5000/5501)
)
```
